News Topic Modeling

### 2017-11-02 paper/R/02_combine.Rmd
new estimation

```{r}
pat <- 'Aus unserem Netzwerk[^"]*'
btw %>%
  filter(grepl(pat, text, perl = TRUE)) %>%
  group_by(site) %>%
  tally(sort = TRUE)
```

@@

### Remove customized stopwords
```{r}
mystopwords <- c("focus","online","spiegel", "stern", "handelsblatt", "de", "bild","bildplus", "zeit", "ersten", "tagesschau", "wdr", "ndr", "ich","sie","passwort","kommentar","wurde","ihr","der","im","artikel","mehr","ihren","foto","e", "uhr","videolänge","dass","mindestens","das","mail","die","schon","neuer abschnitt", "login", "loggen", "inaktiv", "nwmi", "nwnoa", "polizei","beim","dpa","video","quelle","afp","witters","fotogalerie", "registrierter","als","spiegel","vielen","in","es","bitte","dank","unserer","nutzer","sei","beitrag","user","seit","zeichen", "datenschutzerklärung","premium","nutzungsbedingungen","nutzungsrechte","pflichtfelder","registrierung","anzeige","großbuchstaben","sonderzeichen","html","seitennavigation","fullscreen","statista","club","sagte","borenda","spreepicture","shopübersicht")

>>>>

```{r}
btw %>%
  # remove articles that contain a daily overview
  filter(!grepl("News des Tages", title.text)) %>%
  # remove articles that dont contain text
  filter(!grepl('</div>', title.text)) -> btw
```

```{r}
pat <- '</div>'
btw %>%
  filter(grepl(pat, text, perl = TRUE)) %>%
  group_by(site) %>%
  tally(sort = TRUE)
```

@@

### Remove customized stopwords
```{r}
mystopwords <- c("focus","online","spiegel", "stern", "handelsblatt", "de", "bild","bildplus", "zeit", "ersten", "tagesschau", "wdr", "ndr", "ich","sie","passwort","kommentar","wurde","ihr","der","im","artikel","mehr","ihren","foto","e","seien", "uhr","videolänge","dass","mindestens","das","mail","die","schon","neuer abschnitt", "login", "loggen", "inaktiv", "nwmi", "nwnoa", "polizei","beim","dpa","video","quelle","afp","witters","fotogalerie","wurden","worden","wegen","sagt","immer","gibt","geht", "registrierter","als","spiegel","vielen","in","es","bitte","dank","unserer","nutzer","sei","beitrag","user","seit","zeichen", "datenschutzerklärung","premium","nutzungsbedingungen","nutzungsrechte","pflichtfelder","registrierung","anzeige","großbuchstaben","sonderzeichen","html","seitennavigation","fullscreen","statista","club","sagte","borenda","spreepicture","shopübersicht")

### 2017-11-02 paper/R/03_stm.Rmd
new estimation

```{r Setup}
suppressPackageStartupMessages({
  library(dplyr)       # Data manipulation
  library(stringr)     # String manipulation
  library(lubridate)   # Date and time manipulation
  library(purrr)       # Functional programming
  library(tidyr)       # Reshaping
  library(magrittr)    # Advanced piping
  library(pushoverr)   # Pushover notifications
  library(doMC)        # Parallel Computing
  library(readr)       # Importing data
  library(tibble)      # Better data frames

  library(ggplot2)     # Static data visualization
  library(ggrepel)     # Repel text labels
  library(ggiraph)     # GGplot interactive
  library(scales)      # Scales
  library(viridis)     # Viridis color scales
  library(htmlwidgets) # JS visuliaztions
  library(htmltools)   # Arbitrary html
  library(ggjoy)       # Create joyplots
  library(gganimate)   # Animating ggplots
  library(tweenr)      # Tweening charts

@@

```{r caching, echo = FALSE}
load("../output/models/LDAGibbs 40 .Rda")
load("../output/models/STM 40 .Rda")
#load("../output/btw_combined.Rda")
knitr::opts_chunk$set(eval = TRUE, warning = FALSE)

@@

```{r eval=FALSE, include=FALSE}
# Build Corpus
btw %>%
  filter(nchar(text_cleaned) > 5) -> btw

# Build corpus
df.lda <- Corpus(VectorSource(btw$text_cleaned))
# Create document-term matrix
dtm <- DocumentTermMatrix(df.lda)
# Assign name to docs
rownames(dtm) <- make.unique(btw$title)

# ---------- STM ----------- #

btw$month <- month(btw$date)
# Process data
processed <- textProcessor(btw$text_cleaned, metadata = btw[,c("date","month","site","text_cleaned","text")],
                           wordLengths = c(2,Inf),
                           lowercase = F,
                           removestopwords = F,
                           removenumbers = F,
                           removepunctuation = F,
                           stem = F)
out <- prepDocuments(processed$documents, processed$vocab, processed$meta)
```

We then train the LDA on the full dataset, with $k = 40$.

```{r eval=FALSE, include=FALSE}
## Run LDA using Gibbs sampling

#Set parameters for Gibbs sampling (see https://eight2late.wordpress.com/2015/09/29/a-gentle-introduction-to-topic-modeling-using-r/)
burnin <- 1000
iter <- 1000
keep <- 50

k = 40 # Topics
```

```{r}
t1 <- Sys.time()
ldaOut <-LDA(dtm, k, method="Gibbs", control=list(burnin=burnin, iter = iter, keep = keep))
t2 <- Sys.time()
t2 - t1

save(ldaOut, btw, dtm, file=paste("../output/models/LDAGibbs",k,".Rda"))

# ---------- STM ----------- #

@@

# Word-topic probabilities
ldaOut %>% tidy("beta") -> posts.wtp

# Document-topic probabilities
ldaOut %>% tidy("gamma") -> posts.dtp

@@

@@

# The topics function from the package is used to extract the most likely topic for each document 
btw.topics <- topics(ldaOut, 2)

>>>>

```{r Setup}
suppressPackageStartupMessages({
  library(dplyr)       # Data manipulation
  library(stringr)     # String manipulation
  library(lubridate)   # Date and time manipulation
  library(purrr)       # Functional programming
  library(tidyr)       # Reshaping
  library(magrittr)    # Advanced piping
  library(pushoverr)   # Pushover notifications
  library(doMC)        # Parallel Computing
  library(readr)       # Importing data
  library(tibble)      # Better data frames
  library(data.table)

  library(ggplot2)     # Static data visualization
  library(ggrepel)     # Repel text labels
  library(ggiraph)     # GGplot interactive
  library(scales)      # Scales
  library(viridis)     # Viridis color scales
  library(htmlwidgets) # JS visuliaztions
  library(htmltools)   # Arbitrary html
  library(ggjoy)       # Create joyplots
  library(gganimate)   # Animating ggplots
  library(tweenr)      # Tweening charts
  library(igraph)

@@

```{r caching, echo = FALSE}
load("../output/models/STM 40 .Rda")
knitr::opts_chunk$set(eval = TRUE, warning = FALSE)

@@

```{r eval=FALSE, include=FALSE}
# Build Corpus
btw$month <- month(btw$date)
# Process data
processed <- textProcessor(btw$text_cleaned, metadata = btw[,c("date","month","site","text_cleaned","text")],
                           wordLengths = c(2,Inf),
                           lowercase = F,
                           removestopwords = F,
                           removenumbers = F,
                           removepunctuation = F,
                           stem = F)
out <- prepDocuments(processed$documents, processed$vocab, processed$meta)
```

We then train the Structural Topic Model on the full dataset, with $k = 40$.

```{r}
k = 40 # Topics

@@

# Word-topic probabilities
stmOut %>% tidy("beta") %>% filter(!is.na(topic)) -> posts.wtp

# Document-topic probabilities
stmOut %>% tidy("gamma") -> posts.dtp
```

@@

labelTopics(stmOut, n=7)
```

@@

# The topics function from the package is used to extract the most likely topic for each document 
btw.topics <- topics(stmOut, 2)

### 2017-11-02 paper/R/03_topic model.Rmd
new estimation

```{r caching, echo = FALSE}
load("../output/models/LDAGibbs 40 .Rda")
#load("../output/btw_combined.Rda")
knitr::opts_chunk$set(eval = TRUE, warning = FALSE)

btw$articleID <- rownames(btw)
```

@@

```{r eval=FALSE, include=FALSE}
# Build Corpus
btw %>%
  filter(nchar(text_cleaned) > 5) -> btw

# Build corpus
df.lda <- Corpus(VectorSource(btw$text_cleaned))
# Create document-term matrix
dtm <- DocumentTermMatrix(df.lda)
# Assign name to docs
rownames(dtm) <- make.unique(btw$title)
```

@@

doctopics.df %>%
  transmute(title = rownames(.),
            topic = V1) -> doctopics.df

doctopics.df$articleID <- as.integer(rownames(doctopics.df))

# Add Topic to origian DF
topics_mapping.dt <- btw %>%
  mutate(articleID = as.integer(articleID)) %>%
  inner_join(.,doctopics.df, by="articleID") %>%

  group_by(topic) %>%
  slice(1) %>%
  select(topic, title.text, articleID) 
```

@@

set.seed(9272)
posts.dtp %>% 
  group_by(title = document) %>% 
  summarise(topic = min(topic[gamma == max(gamma)]), gamma = max(gamma)) %>% 
  ungroup() %>% 
  inner_join(btw %>% select(title, articleID), by = "title") %>% 
  inner_join(topics_mapping.dt, by = "topic") %>% 
  mutate(topic_title = 
           paste0("Topic ", formatC(topic, flag = "0", width = 2), 
                  " - ", topic_name)) %>% 
  group_by(topic_name) %>% 
  top_n(300, gamma) %>%
  sample_n(5, replace = TRUE) %>% 
  mutate(row = row_number()) %>% 
  ungroup() ->
  posts_classification.sdt

@@

@@

posts.dtp %>% 
  group_by(document) %>% 
  filter(gamma == max(gamma)) %>% 
  mutate(allocation = 1/n()) %>% 
  ungroup() ->
  posts_classification.dt

@@

set.seed(7292)
posts_classification.dt %>% 
  # Add post information
  inner_join(
    btw %>% 
      select(document = title, date, 
             title, articleID) %>% 
      mutate(post_period = 
               date %>% 
               with_tz("Europe/Paris") %>% 
               floor_date("week")),
    by = "document"
  ) %>%
  # Add topic information
  left_join(topics_mapping.dt, by = "topic") %>%
  # Summarise into tidy dataframe
  group_by(post_period, topic_name) %>% 
  summarise(
    articles = sum(allocation),
    sample_articles = 
      paste0("<li>", sample(document, 3, replace = TRUE ), "</li>", collapse = "<br>")
  ) %>% 
  ungroup() %>% 
  mutate(
    tooltip_text = 
      paste0(
        "<b>", topic_name, "</b> in period of <b>", post_period, "</b><br>",
        "<b>Number of Articles:</b> ", comma(round(articles, 0)), "<br>",
        "<b>Sample Articles:</b><br><ul>", sample_articles, "</ul>"
      )
  ) %>% 
  # Mark peaks 
  group_by(topic_name) %>% 
  mutate_at(
    vars(articles), 
    funs(mean = mean(.), sd = sd(.), zscore = (. - mean(.))/sd(.))
  ) %>% 
  ungroup() %>% 
  mutate_at(
    vars(ends_with("zscore")),
    funs(peak = . >	1.96)
  ) %>% 
  mutate(tooltip_text = str_replace_all(tooltip_text, "'", "")) ->
  chart_topic_trends.dt
```

@@

# Prepare dataset for joy plot
posts_classification.dt %>% 
  inner_join(
    btw %>% 
      select(
        document = title, 
        news_page = site, 
        title, 
        date
      ),
    by = "document"
  ) %>% 
  inner_join(topics_mapping.dt, by = "topic") %>% 
  inner_join(
    btw %>% 
      group_by(document = title) %>%
      summarise(reactions = sum(fb_shares)) %>%
      ungroup(),
    by = "document"
  ) %>%
  group_by(
    topic_name, 
    # 8 hour time difference (Manila)
    week = floor_date(date, "week")
  ) %>% 
  summarise(
    num_articles  = sum(allocation),
    num_reactions = sum(reactions * allocation)
  ) %>%
  #mutate(peak = week[num_articles == max(num_articles)]) %>%
  ungroup() %>%
  arrange(week) %>%
  mutate(
    topic_name = factor(topic_name, levels = rev(unique(topic_name))),
    # ID needed for linking back information after tweening frames
    id = as.character(row_number()),
    # Maximum height for ridgeline is 9
    articles_height = num_articles/max(num_articles) * 9,
    reactions_height = num_reactions/max(num_reactions) * 9
  ) ->
  posts_time_trend.dt

@@

set.seed(7292)
posts_classification.dt %>% 
  inner_join(
    btw %>% select(document = title, news_page = site, title),
    by = "document"
  ) %>%
  inner_join(topics_mapping.dt, by = "topic") %>% 
  group_by(news_page, topic_name) %>%
  summarise(
    articles = sum(allocation),
    sample_articles = 
      paste0(
        "<li>", sample(document, pmin(3, length(document))), "</li>",
        collapse = "<br>"
      )
  ) %>%
  ungroup() %>% 
  group_by(news_page) %>% 
  mutate(perc_articles = articles/sum(articles)) %>% 
  ungroup() %>%
  # left_join(
  #   top_100_fbpages.dt %>% select(news_page = page_name, page_title), 
  #   by = "news_page"
  # ) %>% 
  mutate(
    tooltip_text = 
      paste0(
        "<b>", comma(round(articles, 0)), " (", percent(perc_articles), 
        ") </b> of articles from <b>", news_page, 
        "</b> fall under <b>", topic_name, "</b><br><br>Sample Articles:<br><ul>",
        sample_articles, "</ul>"
      ) %>% str_replace_all("'", "")
  ) ->
  news_page_distribution.dt
```

@@

#### Topic Concentration

A common criticism leveled against news media sites is that there is undue emphasis on a particular topic. One way to measure concentration across different topics is the Herfindahl-Hirschman Index (HHI) which is simply $\Sigma_n{s^2}$ where $s$ is the share of each topic.

```{r Topic concentration chart, eval = TRUE}

{
  news_page_distribution.dt %>% 
    group_by(news_page) %>% 
    arrange(desc(articles)) %>% 
    summarise(
      hh_index  = sum(perc_articles ^ 2),
      t3_topics = paste0(topic_name %>% head(3), collapse = "<br>")
    ) %>% 
    ungroup() %>% 
    arrange(hh_index) %>% 
    mutate(news_page = factor(news_page, unique(news_page))) %>% 
    ggplot(aes(x = news_page, y = hh_index * 100)) +
    geom_bar_interactive(
      aes(
        tooltip = 
          paste0("<b>Top Topics:</b><br>", t3_topics) %>% 
          str_replace_all("'", "")
      ),
      stat = "identity", 
      fill = c("#008B45")
    ) +
    geom_text(aes(label = comma(round(hh_index * 100, 1))), hjust = 1.5, 
              color = "white", family = "Roboto") +
    scale_y_log10(
      name = "Herfindahl-Hirschman Index (HHI)\n(100 = most concentrated)",
      expand = c(0, 0)
    ) +
    labs(
      title = "TOPIC CONCENTRATION OF PHILIPPINE NEWS PAGES",
      subtitle = "Herfindahl-Hirschman Index (HHI)",
      caption = "
DATA SOURCE: Facebook Graph API

CHART NOTES:
  1. Hover over each tile to see infomration and to see top topics
  2. Topics were derived through Latent Dirichlet Allocation (LDA) with 40 topics
  3. Articles were classified according to the topic that has the highest probability

TROY JAMES PALANCA | TJPALANCA.COM
      "
    ) +
    coord_flip() +
    theme(
      axis.title.y       = element_blank(),
      panel.grid.major.y = element_blank(),
      plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.subtitle      = element_text(size = 10),
      plot.caption       = element_text(size = 8),
      plot.title         = element_text(size = 16)
    )
} %>% 
  ggiraph(
    ggobj = .,
    width_svg = 10,
    height_svg = 8,
    width = 1,
    tooltip_extra_css = "
    font-family: Roboto; 
    background-color: #000; 
    color: #fff; font-size: 16px;
    padding: 5px;"
  ) %>% 
  prependContent(
    tags$link(
      href = "https://fonts.googleapis.com/css?family=Roboto:400,700", 
      rel = "stylesheet"
    )
  ) -> chart_topic_concentration.wdgt

saveWidget(
  widget = chart_topic_concentration.wdgt,
  file = "../figs/05-topic-concentration.html",
  selfcontained = FALSE,
  libdir = "../figs/js",
  background = "#fafafa"
)

chart_topic_concentration.wdgt

>>>>

```{r caching, echo = FALSE}
#load("../output/models/LDAGibbs 40 .Rda")
load("../output/btw_combined.Rda")
knitr::opts_chunk$set(eval = TRUE, warning = FALSE)

btw$articleID <- as.integer(rownames(btw))
```

@@

```{r eval=FALSE, include=FALSE}
# Build Corpus
df.lda <- Corpus(VectorSource(btw$text_cleaned))
# Create document-term matrix
dtm <- DocumentTermMatrix(df.lda)
```

@@

doctopics.df %>%
  transmute(topic = V1) -> doctopics.df
doctopics.df$articleID <- as.integer(rownames(doctopics.df))

# Add Topic to origian DF
btw %>%
  inner_join(.,doctopics.df, by="articleID") -> btw

topics_mapping.dt <- btw %>%  
  group_by(topic) %>%
  slice(1) %>%
  select(topic,title, title.text, articleID) 
```

@@

set.seed(9272)

posts_classification.sdt <-
posts.dtp %>% 
  group_by(articleID = as.numeric(document)) %>% 
  summarise(topic = min(topic[gamma == max(gamma)]), gamma = max(gamma)) %>% 
  ungroup()  %>%
  inner_join(btw %>% select(title, articleID), by = "articleID") %>%
  inner_join(topics_mapping.dt %>% select(topic, topic_name), by = "topic") %>% 
  mutate(topic_title = 
           paste0("Topic ", formatC(topic, flag = "0", width = 2), 
                  " - ", topic_name)) %>% 
  group_by(topic_name) %>% 
  top_n(300, gamma) %>%
  sample_n(7, replace = TRUE) %>% 
  mutate(row = row_number()) %>% 
  ungroup()

@@

#### Topic-Media Distribution

```{r}
ggsave(
  plot = {
    btw %>%
      ggplot(aes(as.factor(topic), fill =site)) +
      geom_bar() +
      scale_x_discrete(labels = paste0(topicLabel$topic,":",topicLabel$topic_name)) +
      theme(axis.text.x = element_text(size=8, angle=90),
        legend.text = element_text(size = 8)) +
      labs(x = "", fill = "", 
           title    = "Topic Frequency by Media",
        subtitle = "")

  },

  filename = "../figs/01-topic-distribution.png",
  device = "png",
  dpi = 600
    )
```
![Facebook news samples](../figs/01-topic-distribution.png)

@@

posts.dtp %>% 
  group_by(document) %>% 
  filter(gamma == max(gamma)) %>% 
  mutate(allocation = 1/n()) %>% 
  ungroup() %>%
  mutate(document = as.numeric(document)) ->
  posts_classification.dt

@@

set.seed(7292)
posts_classification.dt %>% 
  # Add post information
  inner_join(
    btw %>% 
      select(document = articleID, date, 
             title, articleID) %>% 
      mutate(post_period = 
               date %>% 
               with_tz("Europe/Paris") %>% 
               floor_date("week")),
    by = "document"
  ) %>%
  # Add topic information
  left_join(topics_mapping.dt %>% select(topic, topic_name), by = "topic") %>%
  # Summarise into tidy dataframe
  group_by(post_period, topic_name) %>% 
  summarise(
    articles = sum(allocation),
    sample_articles = 
      paste0("<li>", sample(title, 3, replace = TRUE ), "</li>", collapse = "<br>")
  ) %>% 
  ungroup() %>%
  mutate(
    tooltip_text = 
      paste0(
        "<b>", topic_name, "</b> in period of <b>", post_period, "</b><br>",
        "<b>Number of Articles:</b> ", comma(round(articles, 0)), "<br>",
        "<b>Sample Articles:</b><br><ul>", sample_articles, "</ul>"
      )
  ) %>% 
  # Mark peaks 
  group_by(topic_name) %>% 
  mutate_at(
    vars(articles), 
    funs(mean = mean(.), sd = sd(.), zscore = (. - mean(.))/sd(.))
  ) %>% 
  ungroup() %>% 
  mutate_at(
    vars(ends_with("zscore")),
    funs(peak = . >	1.96)
  ) %>% 
  mutate(tooltip_text = str_replace_all(tooltip_text, "'", "")) ->
  chart_topic_trends.dt
```

@@

# Prepare dataset for joy plot
posts_classification.dt %>% 
  inner_join(
    btw %>% 
      select(
        document = articleID, 
        news_page = site, 
        title, 
        date
      ),
    by = "document"
  ) %>% 
  inner_join(topics_mapping.dt, by = "topic") %>% 
  inner_join(
    btw %>% 
      group_by(document = articleID) %>%
      summarise(reactions = sum(fb_shares)) %>%
      ungroup(),
    by = "document"
  ) %>%
  group_by(
    topic_name, 
    # 8 hour time difference (Manila)
    week = floor_date(date, "week")
  ) %>% 
  summarise(
    num_articles  = sum(allocation),
    num_reactions = sum(reactions * allocation)
  ) %>%
  #mutate(peak = week[num_articles == max(num_articles)]) %>%
  ungroup() %>%
  arrange(week) %>%
  mutate(
    topic_name = factor(topic_name, levels = rev(unique(topic_name))),
    # ID needed for linking back information after tweening frames
    id = as.character(row_number()),
    # Maximum height for ridgeline is 9
    articles_height = num_articles/max(num_articles) * 9,
    reactions_height = num_reactions/max(num_reactions) * 9
  ) ->
  posts_time_trend.dt

@@

set.seed(7292)
posts_classification.dt %>% 
  inner_join(
    btw %>% select(document = articleID, news_page = site, title),
    by = "document"
  ) %>%
  inner_join(topics_mapping.dt, by = "topic") %>% 
  group_by(news_page, topic_name) %>%
  summarise(
    articles = sum(allocation),
    sample_articles = 
      paste0(
        "<li>", sample(document, pmin(3, length(document))), "</li>",
        collapse = "<br>"
      )
  ) %>%
  ungroup() %>% 
  group_by(news_page) %>% 
  mutate(perc_articles = articles/sum(articles)) %>% 
  ungroup() %>%
  # left_join(
  #   top_100_fbpages.dt %>% select(news_page = page_name, page_title), 
  #   by = "news_page"
  # ) %>% 
  mutate(
    tooltip_text = 
      paste0(
        "<b>", comma(round(articles, 0)), " (", percent(perc_articles), 
        ") </b> of articles from <b>", news_page, 
        "</b> fall under <b>", topic_name, "</b><br><br>Sample Articles:<br><ul>",
        sample_articles, "</ul>"
      ) %>% str_replace_all("'", "")
  ) ->
  news_page_distribution.dt
```

@@
